### 2020-07-14 response/response-demography-3.tex

\begin{enumerate}

\item It is not clear why the author restricts the analysis to Hindu
women. Given that data across four rounds of the NFHS data is being
pooled, I think there is enough data to examine this question for the
non-Hindu (and in particular Muslim) women. Given that patterns of son
preference varies across religious groups, I think it is worth examining
the question for non-Hindus. At the very lease include a discussion of
why the non-Hindus have not been included.

[Response:]

I focus on the Hindu population because they are the majority population group 
and because the prior literature argue that this group has a strong son preference and 
consequently uses sex selection.%
\footnote{
Sikhs and Jains likely have even higher use of sex selection than Hindus but Sikhs 
constitutes less than two percent of the population and Jains less than half a percent.
}
Since there is little evidence that Muslim use sex selection, combining Hindu and Muslim 
would lead to biased results.

How birth spacing of Muslim women changed is certainly of interest and is worthy
of a separate analysis.
However, there are two problems with expanding the analysis.
First, although the lastest round of NFHS provides enough data for the analyses, 
it would be hard to provide much in the way of comparison with earlier periods.
This problem becomes especially acute once the Muslim population is split by education 
levels to make the results comparable to the Hindu population.
Second, the paper is already very long and adding even more tables/graphs and the
necessary discussions of the results does not seem feasible.

I have edited the discussion in the ``Data'' section to make the reason for focusing
on Hindus clearer, and hope to write a separate paper/note looking at how spacing
has changed for Muslims later.

### 2020-07-14 paper/sex_selection_spacing-ver4.tex
Combine prediction calculation in Estimation Strategy

Equation (\ref{eq:hazard}) is equivalent to the logistic hazard model and has the same 
likelihood function as the multinomial logit model \citep{allison82,jenkins95}.
Hence, splitting spells into smaller intervals---here equal to three months---and 
treating them as observations, I can estimate the model using a standard multinomial 
logit model. 
In the reorganized data, the outcome variable is 0 if the woman does not have a child 
in a given interval (the base outcome), 1 if she gives birth to a son in that interval, 
and 2 if she gives birth to a daughter in that interval.

I use the model to predict birth interval measures, the probability of having a birth, 
and the sex ratio.%
\footnote{
The interpretation of competing risk model coefficients is challenging \citep{thomas96}.
}
The predicted sex ratio is the weighted average of individual predicted sex ratios, using 
parity progression probabilities as weights.
For an individual woman, the estimated percent of births that are boys at $t$, $\hat{Y_t}$, is:  
\begin{equation}
\hat{Y}_t 
= 
\frac{ P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t )}
{ P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t) + P(g_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t )} 
\times 100,
\label{eq:probability_son}
\end{equation}
where $b$ and $g$ are the predicted probabilities of having a boy and a girl, respectively,
conditional on the explanatory variables and not having had a child before $t$.
Weighting the percentages boys with the likelihood of exiting the spell with a birth across all 
$t$ gives the predicted percent of boys over the entire spell for an individual.%
\footnote{
Imagine $T=2$. 
If 54 and 66 percent of births are boys and the likelihood of giving birth 20 and 40 percent, 
then the predicted sex ratio is $\frac{54*0.2+66*0.4}{0.2+0.4} = 62$ percent boys. 
}

@@

I use four periods: 1972--1984, 1985--1994, 1995--2004, and 2005--2016.
The first covers before sex selection became available and the second from when sex 
selection became available until the Prenatal Diagnostic Techniques Act.
I have split the period from 1995 until 2016 into two to examine if there is support for 
the prior literature's hypothesized reversal in child sex ratios and son 
preference in India.
The allocation of spells into periods is determined by when conception, and, therefore, 
decisions on sex selection, can begin. 
Hence, some spells cover two periods, which may bias downward the differences between the 
periods.

@@

% Explain question and what I am doing
The first question I address is how birth spacing responded to the significant changes, 
especially the spread of sex selection, that India experienced. 
I estimate the competing risk hazard model for each spell, education, and period 
subsample and then predict average birth interval in months, sex ratio, and the 
probability of having a birth in that spell.

Figures \ref{fig:bootstrap_low} through \ref{fig:bootstrap_highest}
show the predicted outcomes for the four education levels by area of residence.%
\footnote{
For legibility, the graphs do not show standard errors.
The graphed values, with standard errors, are shown in Appendix Tables 
\ref{tab:avg_sex_ratio_low} through \ref{tab:avg_sex_ratio_highest}.
Furthermore, I do not show the results for the fourth spell the best-educated women since 
the low number of births, especially with one or more boys, make the spacing and sex ratio 
results very noisy.
Results are available upon request.
}

The predicted parity progression probability is the likelihood of giving birth by the
imposed censoring based on standard survival curve calculations.%
\footnote{
Most sterilizations take place right after a woman gives birth, and,
therefore, do not show up in the samples used. 
Furthermore, sterilization depends strongly on the sex composition of prior children
with lower probabilities, the fewer boys. 
The effect is to bias downward the differences in parity progression probabilities.
}
The predicted sex ratio captures the percent of births that are boys for women in 
the sample when childbearing for that spell is over, calculated as shown above.
For comparison, I also show the natural sex ratio, approximately 51.2 percent 
\citep{ben-porath76b,jacobsen99,Portner2015b}.
To find the expected average birth interval, I first calculate, for each woman, 
the probability of giving birth in each period $t$, which I use as weights to calculated 
her expected spell lengths. 
I then average the individual expected spell lengths across women using their parity 
progression probabilities as weights.
Finally, I add nine months to account for spell start to arrive at birth interval.

@@

Figures \ref{fig:combined_p25_p75_urban} and \ref{fig:combined_p25_p75_rural} 
show how the predicted 25th and 75th percentile birth intervals have changed over time 
by spell and education.
For each woman in a given spell/period combination, I calculate the time point at 
which there is a given percent chance that she will have given birth, relative to the
probability that she will eventually give birth in that spell.
The reported statistics is the average of the percentile duration
across all women in a given sample using the individual predicted
probabilities of having had a birth by the end of the spell as weights.%
\footnote{
The Appendix shows conditional survival curves for all groups, and 
Appendix Tables \ref{tab:p25_p50_p75_low} through \ref{tab:p25_p50_p75_highest} 
show 25th, 50th, and 75th percentiles durations with bootstrapped standard errors.
}

>>>>

Equation (\ref{eq:hazard}) is equivalent to the logistic hazard model and has the same 
likelihood function as the multinomial logit model \citep{allison82,jenkins95}.
Hence, splitting spells into smaller intervals---here equal to three months---and 
treating them as observations, I can estimate the model using a standard multinomial 
Logit model. 
In the reorganized data, the outcome variable is 0 if the woman does not have a child 
in a given interval (the base outcome), 1 if she gives birth to a son in that interval, 
and 2 if she gives birth to a daughter in that interval.

I use the model to predict birth interval measures, parity progression probabilities, 
and the sex ratio rather than present coefficients since the interpretation 
of competing risk model coefficients is challenging \citep{thomas96}.
The predicted parity progression probability is the likelihood of giving birth by the
imposed censoring based on standard survival curve calculations averaged across all
women in a given sample.

For birth intervals, I estimate the average birth interval and a set of percentile
birth intervals.
To find the average birth interval, I calculate, for each woman, the probability of 
giving birth in each $t$, and her expected spell length from these probabilities.
I then average the individual expected spell lengths across women using their parity 
progression probabilities as weights.
Finally, I add nine months because spells begin nine months after the previous birth.
The process is similar for the percentile birth intervals, except that I use the time 
point when there is a given percent probability that a woman will give birth, conditional 
on her giving birth in the spell.

The predicted sex ratio is the weighted average of individual predicted sex ratios, using 
parity progression probabilities as weights.
For an individual woman, the estimated percent of births that are boys at $t$, $\hat{Y_t}$, is:  
\begin{equation}
\hat{Y}_t 
= 
\frac{ P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t )}
{ P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t) + P(g_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t )} 
\times 100,
\label{eq:probability_son}
\end{equation}
where $b$ and $g$ are the predicted probabilities of having a boy and a girl, respectively,
conditional on the explanatory variables and not having had a child before $t$.
Weighting the percentages boys with the likelihood of exiting the spell with a birth across all 
$t$ gives the predicted percent of boys over the entire spell for an individual.%
\footnote{
Imagine $T=2$. 
If 54 and 66 percent of births are boys and the likelihood of giving birth 20 and 40 percent, 
then the predicted sex ratio is $\frac{54*0.2+66*0.4}{0.2+0.4} = 62$ percent boys. 
}

@@

I use four periods: 1972--1984, 1985--1994, 1995--2004, and 2005--2016.
The first covers before sex selection became available and the second from when sex 
selection became available until the Prenatal Diagnostic Techniques Act.
I have split the period from 1995 until 2016 into two to examine if there is support for 
the prior literature's hypothesized reversal in child sex ratios and son 
preference in India.

The allocation of spells into periods is determined by when conception, and, therefore, 
decisions on sex selection, can begin. 
Hence, some spells cover two periods, which may bias downward the differences between the 
periods.
Most sterilizations take place right after a woman gives birth, and,
therefore, do not show up in the samples used. 
Furthermore, sterilization depends strongly on the sex composition of prior children
with lower probabilities, the fewer boys. 
The effect is to bias downward the differences in parity progression probabilities.

@@

% Explain question and what I am doing
The first question I address is how birth spacing responded to the significant changes, 
especially the spread of sex selection, that India experienced. 
I estimate the competing risk hazard model for each spell, education, and period 
combination and then predict average birth interval in months, sex ratio, and the 
probability of having a birth in that spell.

Figures \ref{fig:bootstrap_low} through \ref{fig:bootstrap_highest}
show the predicted outcomes for the four education levels by area of residence.%
\footnote{
For legibility, the graphs do not show standard errors.
The graphed values, with standard errors, are shown in Appendix Tables 
\ref{tab:avg_sex_ratio_low} through \ref{tab:avg_sex_ratio_highest}.
Furthermore, I do not show the results for the fourth spell the best-educated women since 
the low number of births, especially with one or more boys, make the spacing and sex ratio 
results very noisy.
Results are available upon request.
}
For comparison, the graphs for the sex ratio also show the natural sex ratio, 
approximately 51.2 percent \citep{ben-porath76b,jacobsen99,Portner2015b}.

@@

Figures \ref{fig:combined_p25_p75_urban} and \ref{fig:combined_p25_p75_rural} 
show how the predicted 25th and 75th percentile birth intervals have changed over time 
by spell and education.%
\footnote{
The Appendix shows conditional survival curves for all groups, and 
Appendix Tables \ref{tab:p25_p50_p75_low} through \ref{tab:p25_p50_p75_highest} 
show 25th, 50th, and 75th percentiles durations with bootstrapped standard errors.
}
